Notes for Week 6/9/19 - 6/11/19

### 2020-08-19 notes/weekly_notes/3.5.20.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

For an arbitrary graph $G$ on $n$ nodes with graph Laplacian $L_G = D_G - A_G$, let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $i$th eigenvalues $\lambda_k(G)$. The graph Laplacian eigenvector projection test is a linear projection test; letting $\kappa$ be some integer between $1$ and $n$, we define $\Pi_{k,G}(\theta) := \frac{1}{\sqrt{n}}\sum_{k = 1}^{\kappa} \dotp{v_k}{\theta}$ to be a suitable rescaling of the projection of $\theta$ onto the span of $v_1(G),\ldots,v_k(G)$.

@@

We let the space 
\begin{equation*}
\wt{H}^{s}([0,1]^d;L) = \Bigl\{ \sum_{k \in \mathbb{N}^d} \theta_k \varphi_k: \sum_{k \in \mathbb{N}^d} \theta_k^2 a_k^{2} < L \Bigr\}
\end{equation*}
where in the one-dimensional case,
\begin{equation*}
a_{k - 1} =
\begin{cases*}
(k - 1)^s,~~ \textrm{if $k - 1$ is even} \\
(k - 2)^s,~~ \textrm{if $k - 1$ is odd}
\end{cases*}
\end{equation*}
and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$. It is not hard to show that $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some boundary conditions
\begin{equation*}
\wt{H}^s([0,1]^d;L) = \Bigl\{f \in H^{s}([0,1]^d;L'): \forall x_0 \in \partial [0,1]^d, \frac{\partial^{\ell} f}{\partial \mathbf{n}^{\ell}}(x_0) = 0,~~\textrm{for $\ell = 1,\ldots,s-1$} \Bigr \} \supseteq H_0^s([0,1]^d;L')
\end{equation*}
for a constant $L'$ which depends only on $L$ and $s'$; we work through the details for the case $d = 1$ in Section~\ref{subsec:sobolev_class_equivalence}.

@@

We extend the results of Lemma~\ref{lem:grid_sobolev_approximation_error_5} to hold for all $d \geq 1$ in Lemma~\ref{lem:grid_sobolev_approximation_error_6}. Although Lemma~\ref{lem:grid_sobolev_approximation_error_6} strictly dominates Lemma~\ref{lem:grid_sobolev_approximation_error_5}, we keep the latter around since its proof is somewhat easier to follow. 
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_6}
	Suppose $f = \sum_{k \in \Nbb^d} \theta_k \varphi_k$ satisfies $f \in \wt{H}^s([0,1]^d;L)$ for some $2s > d$. Then, there exists a constant $c$ which depends only on $s$ and $d$ such that
	\begin{equation}
	\label{eqn:grid_sobolev_approximation_error_6}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s/d} + \sum_{k \in \Nbb \setminus [K]^d} \theta_k^2
	\end{equation}
\end{lemma}

@@

\textcolor{red}{Work through details.}

\end{document}

>>>>

For an arbitrary graph $G$ on $n$ nodes with graph Laplacian $L_G = D_G - A_G$, let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $i$th eigenvalues $\lambda_k(G)$. The graph Laplacian eigenvector projection test is a linear projection test; letting $\kappa$ be some integer between $1$ and $n$, we define $\Pi_{k,G}(\theta) := \sum_{k = 1}^{\kappa} \dotp{v_k}{\theta}$ to be the projection of $\theta$ onto the span of $v_1(G),\ldots,v_k(G)$.

@@

We let the space 
\begin{equation*}
\wt{H}^{s}([0,1]^d;L) = \Bigl\{ \sum_{k \in \mathbb{N}^d} \theta_k \varphi_k: \sum_{k \in \mathbb{N}^d} \theta_k^2 a_k^{2} < L \Bigr\}
\end{equation*}
where in the one-dimensional case,
\begin{equation*}
a_{k} = (k - 1)^s
\end{equation*}
and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$. It is not hard to show that, when $s$ is an integer, $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some Neumann-type boundary conditions. For example when $d = 1$,
\begin{equation*}
\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L \pi^{s}): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for all odd integers $0 < \ell < s$} \Bigr \} \supseteq H_0^s([0,1];L'),
\end{equation*}
a fact which we prove in Section~\ref{subsec:sobolev_class_equivalence}.

@@

We extend the results of Lemma~\ref{lem:grid_sobolev_approximation_error_5} to hold for all $d \geq 1$ in Lemma~\ref{lem:grid_sobolev_approximation_error_6}. Although Lemma~\ref{lem:grid_sobolev_approximation_error_6} strictly dominates Lemma~\ref{lem:grid_sobolev_approximation_error_5}, we keep the latter around since its proof is somewhat easier to follow. 
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_6}
	Suppose $f = \sum_{k \in \Nbb^d} \theta_k \varphi_k$ satisfies $f \in \wt{H}^s([0,1]^d;L)$ for some $2s > d$. Then, there exists a constant $c$ which depends only on $s$ and $d$ such that
	\begin{equation}
	\label{eqn:grid_sobolev_approximation_error_6}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s/d} + \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2
	\end{equation}
\end{lemma}

@@

We will show that when $s \geq 1$,
\begin{equation*}
\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L'): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for each $\ell < s$ odd.} \Bigr \}
\end{equation*}

First, we show the $\subseteq$ direction. Let $f = \sum_{k = 1}^{\infty} \theta_k \varphi_k$ satisfy $\sum_{k = 1}^{\infty} \theta_k^2 a_k^2 \leq L^2$. Put
\begin{equation*}
f_{\diamond} := \sqrt{2} \sum_{k = 1}^{\infty} \theta_k \phi_k
\end{equation*}
where the functions $\phi_k = \cos(\pi[k - 1]x)$ are orthonormal in $\Leb^2([-1,1])$. Since $\sum_{k = 1}^{\infty} (\sqrt{2}\theta_k)^2 a_k^2 \leq 2L^2$, we know $f_{\diamond} \in H_{\mathrm{per}}^s([-1,1];2L)$. Moreover since $f_{\diamond}$ is an even function, we have that
\begin{equation*}
\sum_{\ell = 0}^{s}\int_{0}^{1} \bigl(f_{\diamond}^{(\ell)}(x)\bigr)^2 \,dx = \frac{1}{2} \sum_{\ell = 0}^{s}\int_{-1}^{1} \bigl(f_{\diamond}^{(\ell)}(x)\bigr)^2 \,dx \leq \frac{1}{\pi^{2s}}L^2
\end{equation*}
where the latter inequality follows from \textcolor{red}{Tsybakov}.
As $f(x) = f_{\diamond}(x)$ for all $x \in [0,1]$, we have established that $f \in H^s([0,1],L')$ for $L' = L/\pi^s$. As for the boundary cases $x = 0$ and $x = 1$, putting
\begin{equation*}
f_N := \sum_{k = 1}^{N} \theta_k \varphi_k
\end{equation*}
we observe the following facts: first that $f_N^{(\ell)}(0) = f_N^{(\ell)}(1) = 0$ for all $N \in \Nbb$ and $\ell$ odd, and second that $f_N^{(\ell)}$ converges uniformly for all $0 \leq \ell < s$. Therefore we have $f^{(\ell)}(x) = \lim_{N \to \infty} f_N^{(\ell)}(x)$ for all $x \in [0,1]$ and $0 \leq \ell < s$, and in particular $f^{(\ell)}(0) = f^{(\ell)}(1) = 0$ for all $0 < \ell < s$ odd.

Now for the $\supseteq$ direction. Let $f \in H^s([0,1];L)$ satisfy $f^{(\ell)}(0)$ for all $\ell = 1,\ldots,s - 1$ odd. Put
\begin{equation*}
f_{\diamond} = f(\abs{x})~~\textrm{for $x \in [-1,1]$}
\end{equation*}
and note the following facts: first that $f_{\diamond}^{(\ell)}(0)$ is well-defined for each $\ell = 0,\ldots, s - 1$, and second that $f_{\diamond}^{(s - 1)}$ is absolutely continuous on $[-1,1]$. As a result, $f_{\diamond}$ is $s$-times weakly differentiable (and $s - 1$ times classically differentiable), and by construction satisfies
\begin{equation*}
\norm{f_{\diamond}}_{H^s([-1,1])} = \sqrt{2} \norm{f}_{H^s([0,1])} \leq 2L
\end{equation*} 
Since clearly $f_{\diamond}^{(\ell)}(-1) = f_{\diamond}^{(\ell)}(1)$ for each $\ell = 0,\ldots, s - 1$, we have that $f_{\diamond} \in H_{\mathrm{per}}^s([-1,1];\sqrt{2}L)$. Since $f_{\diamond}$ is even, it may be expressed as $f_{\diamond} = \sum_{k = 1}^{\infty} \theta_k \phi_k$ where $\sum_{k = 1}^{\infty} \theta_k^2 a_k^{2} \leq 2L^2/\pi^{2s}$. Since $f(x) = f_{\diamond}(x)$ for $x \in [0,1]$, this implies
\begin{equation*}
f(x) = \sum_{k = 1}^{\infty} \theta_k \phi_k(x) = \sum_{k = 1}^{\infty} \frac{1}{\sqrt{2}}\theta_k \varphi_k(x) 
\end{equation*}
and the claim is proved.
\end{document}

### 2020-08-19 notes/weekly_notes/6.18.19.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\begin{theorem}

\end{theorem}

@@

\begin{lemma}
	\label{lem: var_mmd}
	For any $h,g \in \mathcal{H}_1^{d}(L)$,
	\begin{equation*}
	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d
	\end{equation*}
\end{lemma}
\begin{proof}
	We begin by computing the second moment $\Ebb T_{\eta}^2$. Expanding the quadratic form, we obtain
	\begin{align}
	\Ebb T_{\eta}^2 =  \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \Ebb(\ell_i \ell_{i'} \ell_{j} \ell_{j'} A_{ii'} A_{jj'})
	\end{align}

@@

	The following upper bound on the third term is a straightforward consequence of the fact $\eta_r^2 = \eta_r$, and therefore
	\begin{equation}
	\label{eqn: var_mmd_3}
	\Ebb(\eta_r(X,Y)^2) = \Ebb(\eta_r(X,Y)) = r^d. 
	\end{equation}

@@

	Combining \eqref{eqn: var_mmd_1}-\eqref{eqn: var_mmd_4}, we obtain
	\begin{equation*}
	\Var(T_\eta) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d.
	\end{equation*}
\end{proof}

>>>>

We show that a simple test based on the statistic $T_{\eta}$, with proper choice of the bandwidth parameter $r$, achieves minimax optimality over the class of Lipschitz functions $\mathcal{H}_{1}^{d}(L)$.

\begin{theorem}
	\label{thm: mmd_minimax_optimal}
	Fix $r = n^{-2/(4+d)}$, and consider the test 
	\begin{equation*}
	\phi_{\eta,\tau} = \1\left\{ T_{\eta} \geq \tau \right\}, \quad \tau = a n^{-d/(4+d)}
	\end{equation*}
	with $a \geq 1$. There are constants $c_1$ depending only on $p_{\max}$, and $c_2$ depending only on $(L,p_{\max},p_{\min})$, such that
	\begin{equation*}
	R_{\epsilon}^{n}(\phi_{\eta,\tau}; \mathcal{H}_{1}^{d}(L)) \leq c_1/a^2, \quad \text{if}~ \epsilon \geq c_2 a n^{-2/(4 + d)}
	\end{equation*}
\end{theorem}
\begin{proof}
	Our first objective is to show that, for an appropriate choice of $c_1$, $\Ebb_{f,f}^{n}(\phi_{\eta,\tau}) \leq \frac{c_1}{2a^2}$ for all $f \in \mathcal{H}_{1}^{d}(L)$. This is equivalent to proving that
	\begin{equation}
	\label{eqn: mmd_minimax_optimal_2}
	\Pbb \left(T_{\eta} \geq \tau \right) \leq c_1/2a^2
	\end{equation}
	which we now proceed to show.

	From Lemma \ref{lem: mmd_expectation}, we have that $\Ebb_{f,f}(T_{\eta}) = 0$. Therefore, by Chebyshev's inequality, 
	\begin{equation}
	\label{eqn: mmd_minimax_optimal_1}
	\Pbb \left(T_{\eta} \geq \tau \right) \leq \Var(T_{\eta})/\tau^2
	\end{equation}
	In Lemma \ref{lem: var_mmd}, we have upper bounded $\Var(T_{\eta})$ as function of $r$ and $\delta^2(f,g)$. Under the null hypothesis $f = g$, $\delta^2(f,g) = 0$, and along with our choice of $r = n^{-2/(d + 4)}$, the upper bound on $\Var(T_{\eta})$ reduces to
	\begin{equation*}
	\Var(T_{\eta}) \leq \frac{p_{\max}^2}{2} n^{-2d/(d + 4)}
	\end{equation*}
	By our choice of $\tau = a n^{-d/(4 + d)}$, we therefore have
	\begin{equation*}
	\Var(T_{\eta})/\tau^2 \leq \frac{p_{\max}^2}{2a^2}
	\end{equation*}
	and therefore \eqref{eqn: mmd_minimax_optimal_2} holds with $c_1 := p_{\max}^2$.

	We turn now to bound the type II error, $\Ebb_{f,g}^{(n)}(\phi_{\eta,\tau})$, for all $f,g \in \mathcal{H}_{1}^{d}(L)$ such that $\delta^2(f,g) \geq \epsilon = c_2 n^{-2/(4+d)}$. Our goal now will be to show
	\begin{equation*}
	\Pbb \left(T_{\eta} \leq \tau \right) \leq c_1/2a^2.
	\end{equation*}
	By Chebyshev's inequality we obtain
	\begin{equation}
	\label{eqn: mmd_minimax_optimal_4}
	\Pbb(T_{\eta} \leq \tau) \leq \frac{\Var(T_{\eta})}{(\Ebb_{f,g}(T_{\eta}) - \tau)^2}
	\end{equation}
	We obtain the following lower bound on $\Var(T_{\eta})$ in Lemma \ref{lem: var_mmd} in terms of $r$ and $\delta(f,g)$:
	\begin{equation*}
	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d
	\end{equation*}
	We will split our analysis into cases, based on which term on the right hand side of the preceding equation dominates. \textcolor{red}{TODO: Complete this proof.}

	\paragraph{Case 1, 1st term dominates:}
	We assume
	\begin{equation*}
	2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) > \frac{p_{\max}^2}{2}r^d,
	\end{equation*}
	so that the variance is upper bounded
	\begin{equation*}
	\Var(T_{\eta}) \leq 4 n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right).
	\end{equation*}
	Additionally, in the following case, we will show that for a sufficiently large choice of $c_2$, depending on $p_{\max}, p_{\min}, L$ and $d$, we obtain
	\begin{equation*}
	\Ebb_{f,g} T_{\eta} \geq 2 \tau
	\end{equation*}
	so that 
	\begin{equation*}
	(\Ebb_{f,g} T_{\eta} - \tau)^2 \geq \left(\Ebb_{f,g} T_{\eta}\right)^2/4.
	\end{equation*}
	In Lemma \ref{lem: mmd_expectation}, moreover, we prove the following lower bound on $\Ebb_{f,g} T_{\eta}$ as a function of $r$ and $\delta(f,g)$:
	\begin{equation*}
	\Ebb_{f,g} T_{\eta} \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right).
	\end{equation*}
	We can therefore upper bound $\Var(T_{\eta})/(\Ebb_{f,g}T_{\eta} - \tau)^2$,
	\begin{align*}
	\Var(T_{\eta})/(\Ebb_{f,g}T_{\eta} - \tau)^2 \leq \frac{p_{\max}^6}{p_{\min}^6}  \frac{16 n r^{2d} \delta(f,g) \left(\delta(f,g) + Lr\right)}{(n-1)^2 r^{2d} \delta^2(f,g) \left(\delta(f,g){2^{-d}} - Lr\right)^2}
	\end{align*}

	\paragraph{Case 2, 2nd term dominates:}
	We first lower bound the denominator, and then upper bound the numerator. In Lemma \ref{lem: mmd_expectation}, we prove the following lower bound on $\Ebb T_{\eta}$ when $f,g \in \mathcal{H}_1^{d}(L)$, stated as a function of $r$ and $\delta^2(f,g)$:
	\begin{equation}
	\label{eqn: mmd_minimax_optimal_3}
	\Ebb_{f,g} T_{\eta} \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right).
	\end{equation}
	Given the choice of bandwidth parameter $r = n^{-2/(4+d)}$ and the lower bound $\delta(f,g) \geq c_2 a n^{-2/(d + 4)}$, we may further simplify the expression on the right hand side of \eqref{eqn: mmd_minimax_optimal_3},
	\begin{align*}
	\frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right) & = \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} n^{-2d/(d+4)} \delta(f,g) \left(\frac{\delta(f,g)}{2^d} - Ln^{-2/(d + 4)} \right) \\
	& \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} c_2 a n^{-(2d + 2)/(d+4)} \left(\frac{c_2 a n^{-2/(d + 4)}}{2^d} - Ln^{-2/(d + 4)} \right) \\
	& \geq 2 a^2 n^{1 - (2d + 2)/(d + 4) - 2/(d + 4)} = 2 a^2 n^{-d/(d + 4)}.
	\end{align*}
	where the last inequality follows when $c_2$ is chosen to be sufficiently large relative to $p_{\min}, p_{\max}, L$ and $d$. Given our choice $\tau = a n^{-d/(d + 4)}$, the denominator of \eqref{eqn: mmd_minimax_optimal_4} can therefore be lower bounded
	\begin{equation*}
	(\Ebb_{f,g}(T_{\eta}) - \tau)^2 \geq (2a^2 - a)^2 n^{-2d/d(+4)} \geq a^4 n^{-2d/d(+4)} 
	\end{equation*}
	where the last inequality follows from the fact $2a^2 - a \geq a^2$ for all $a \geq 1$.

	We turn now to upper bounding the numerator of \eqref{eqn: mmd_minimax_optimal_4}. In Lemma \ref{lem: var_mmd} we give the following upper bound on $\Var_{f,g} T_\eta$ as a function of $r$ and $\delta(f,g)$, 
	\begin{equation*}
	\Var_{f,g} T_\eta \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d 
	\end{equation*}
	Similar to before, we apply our choice of $r$ and lower bound on $\delta(f,g)$ and obtain
	\begin{equation*}
	2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d \leq 
	\end{equation*}
\end{proof}

@@

\begin{lemma}
	\label{lem: var_mmd}
	For any $h,g \in \mathcal{H}_1^{d}(L)$,
	\begin{equation*}
	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d
	\end{equation*}
\end{lemma}
\begin{proof}
	We begin by computing the second moment $\Ebb T_{\eta}^2$. Expanding the quadratic form, we obtain
	\begin{align}
	\Ebb T_{\eta}^2 =  \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \Ebb(\ell_i \ell_{i'} \ell_{j} \ell_{j'} A_{ii'} A_{jj'})
	\end{align}

@@

	The following upper bound on the third term is a straightforward consequence of the fact $\eta_r^2 = \eta_r$, and therefore
	\begin{equation}
	\label{eqn: var_mmd_3}
	\Ebb(\eta_r(X,Y)^2) = \Ebb(\eta_r(X,Y)) \leq \frac{p_{\max}^2}{4} r^d. 
	\end{equation}

@@

	Combining \eqref{eqn: var_mmd_1}-\eqref{eqn: var_mmd_4}, we obtain
	\begin{equation*}
	\Var(T_\eta) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d.
	\end{equation*}
\end{proof}

### 2020-08-19 notes/weekly_notes/6.9.19.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}

@@

>>>>

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\wt}[1]{\widetilde{#1}}

@@

\subsection{Relating the eigenvectors to a grid.}

For $\kappa = n^{1/d}$, let $\widetilde{X} = \set{\wt{x} \in [\kappa]^d/\kappa}$ be a grid over the unit cube. Consider the lattice graph $\mathrm{Grid} = (\widetilde{V}, \widetilde{E})$, with vertices $\widetilde{V} = \set{v: v \in [\kappa]^d}$, and edges $\widetilde{E} = \set{(u,v):\widetilde{u}, \wt{v} \in \wt{V}, \norm{\wt{u} - \wt{v}}_1 = 1}$, and write $\widetilde{L} = \widetilde{V} \Lambda \wt{V}^T$ for the eigendecomposition of the associated Laplacian matrix.

We wish to exhibit a mapping $P: \Reals^V \to \Reals^{\wt{V}}$ such that the difference between the statistics $T_{LE}^{(k)} = \ell^T V_k V_k^T \ell$ and $\wt{T}_{LE}^{(k)} = (P \ell)^T \wt{V}_k \wt{V}_k^T (P \ell)$ is small, which we accomplish through the use of optimal transport theory.

\begin{lemma}
	Fix $0 < \delta < 1$. For some number $c_1 > 0$, dependent on $f + g$, $d$, and $\delta$, but not $n$, with probability at least $1 - \delta$ there exists a mapping $T: X \to \wt{X}$ such that
	\begin{equation}
	\label{eqn: optimal_transport_to_grid}
	\norm{X - TX}_{\infty} \leq c_1 \left(\frac{\log(n)}{n}\right)^{1/d}
	\end{equation}
	If $r = c_1 (\log n / n)^{1/d}$ and \ref{eqn: optimal_transport_to_grid} holds, then letting $P(v)$ be the index of $T(x_v)$ on the grid graph, we have for every $\theta \in \Reals^V$,
	\begin{equation}
	(P\theta)^T \wt{L} P \theta \leq \theta^T L \theta \leq (\log n)^{2 + 1/d} (P\theta)^T \wt{L} P \theta.
	\end{equation}
\end{lemma}
